### 2018-06-25 latex/egsr17-cgf/previous_work.tex
applied camera ready fixes

Light transport in participating media is governed by the radiative transfer equation~(RTE), first studied in the context of astrophysics by Chandrashekar~\cite{Chandrasekhar60} and later introduced to computer graphics by Kajiya~\cite{Kajiya86}. Today, this equation is typically solved using Monte Carlo methods. In strongly scattering media or in the presence of highly anisotropic phase functions, these methods can become prohibitively expensive. Consider a high albedo medium like milk, where tracing paths with many scattering events is necessary.

@@

The $P_N$-method has a long history in other fields and was never applied in graphics. Kajia~\cite{Kajiya84} explained the theory, but did not give any details on implementation or how to solve it. In fact, as Max~\cite{Max95} pointed out, it is not clear if Kajiya succeeded at all at applying the method, as all of the results in his paper were produced with a simpler method. This is further strengthened by our observation, that a straight forward finite difference discretization of the $P_N$-equations produces unuseable results, due to oscillation artifacts in the solution. We derive the real-valued $P_N$-equations with a staggered-grid solver, that produces artifact free solutions.

@@

CDA suffers from severe energy loss close to regions with strong density gradients. The problem can be addressed by a modification known as Variable Eddington factor (VEF), which nonlinearly adjusts the diffusion coefficient to improve the solution near density gradients and low-density regions. Flux-limited diffusion, developed in the context of astrophysics by Levermore et al.~\cite{Levermore81} and later introduced to graphics by Koerner et al.~\cite{Koerner14}, is the most prominent example.

VEF is based on the observation, that the behavior of volumetric transport is closely linked to the density of the material: In the diffusive regime, thick highly scattering media is present and causes photons to change directions in short succession, while the transport regime in thin and highly absorbing media causes photons to travel long straight lines. In the absence of scattering (e.g. pure vacuum), the non-linear diffusion coefficient turns the diffusion equation into an advection equation. The idea behind VEF is to derive a better diffusion coefficient by seeing it as a means to spatially blend between the diffusive and pure transport regime.

It is an open and unresolved question, whether the $P_N$-method with truncation at higher order, gives any benefit over using first order non-linear diffusion. This question has also been raised in other domains~\cite{Olson00} and resolving it is one of our motivations.

>>>>

Light transport in participating media is governed by the radiative transfer equation~(RTE), first studied in the context of astrophysics by Chandrasekhar~\cite{Chandrasekhar60} and later introduced to computer graphics by Kajiya~\cite{Kajiya86}. Today, this equation is typically solved using Monte Carlo methods. In strongly scattering media or in the presence of highly anisotropic phase functions, these methods can become prohibitively expensive. For example in the case of a high albedo medium like milk where tracing paths with many a huge number of scattering events is necessary.

%Consider a high albedo medium like milk, where tracing paths with many scattering events is necessary.

@@

The $P_N$-method has a long history in other fields and was never applied in graphics. Kajiya~\cite{Kajiya84} explained the theory, but did not give any details on implementation or how to solve it. In fact, as Max~\cite{Max95} pointed out, it is not clear if Kajiya succeeded at all at applying the method, as all of the results in his paper were produced with a simpler method. This is further strengthened by our observation that a straightforward finite difference discretization of the $P_N$-equations produces unusable results, due to oscillation artifacts in the solution. We derive the real-valued $P_N$-equations with a staggered-grid solver, that produces artifact-free solutions.

@@

CDA suffers from severe energy loss close to regions with strong density gradients. The problem can be addressed by a modification known as the Variable Eddington factor (VEF) method, which nonlinearly adjusts the diffusion coefficient to improve the solution near density gradients and low-density regions. Flux-limited diffusion, developed in the context of astrophysics by Levermore et al.~\cite{Levermore81} and later introduced to graphics by Koerner et al.~\cite{Koerner14}, is the most prominent example.

VEF is based on the observation that the behavior of volumetric transport is closely linked to the density of the material: In the diffusive regime, thick highly scattering media is present and causes photons to change directions in short succession, while the transport regime in thin and highly absorbing media causes photons to travel long straight lines. In the absence of scattering (e.g. pure vacuum), the non-linear diffusion coefficient turns the diffusion equation into an advection equation. The idea behind VEF is to derive a better diffusion coefficient by seeing it as a means to spatially blend between the diffusive and pure transport regime.

It is an open and unresolved question whether the $P_N$-method with truncation at higher order, gives any benefit over using first order non-linear diffusion. This question has also been raised in other domains~\cite{Olson00} and resolving it is one of our motivations.

### 2018-06-25 latex/egsr17-cgf/section_conclusion.tex
applied camera ready fixes

In this paper, we introduced the $P_N$-method to the toolbox of deterministic methods for rendering participating media in computer graphics. We derived and presented the real-valued $P_N$-equations, along with a staggered grid solver for solving them. We showed, how to use the results in a rendering system and ran our solver for various standard problems for validation.

@@

The literature in other fields often states, that the $P_N$ method---when solved in normal form like we do---is able to deal with vacuum regions. We found this misleading. The $P_N$-method in normal form indeed does not break down completely in the presence of vacuum as diffusion based methods do (due to $\sigma_t^{-1}$ in the diffusion coefficient). However, in the presence of vacuum, the condition number of the system matrix becomes infinite and does not converge either. Therefore $P_N$ based methods also require minimum thresholding of the extinction coefficient and offer no benefit for vacuum regions.

@@

Unique to our system is that it uses a computer algebra representation of the equation to solve as input. Discretization in angular and spatial domain is done using manipulation passes on the representation. The stencil code, which is used to populate the system of linear equations, is generated from the expression tree. This way, we can easily deal with coupled PDE's and avoid the time consuming and error prone process of writing stencil code by hand.

When researching the application of $P_N$ in other fields, we came across a rich variety of variations on the $P_N$-method, such as simplified $P_N$ ($SP_N$)~\cite{Ryan10}, filtered $P_N$ ($FP_N$)~\cite{Radice13}, diffusion-correction $P_N$ ($DP_N$)~\cite{Schaefer11} and least-squares $P_N$ ($LSP_N$)~\cite{Hansen14}. These variations have been introduced to address certain problems of the standard $P_N$-method, such as ringing artifacts, dealing with vacuum regions and general convergence. Our solver can be applied to any (potentially coupled) PDE and therefore can generate stencil code for all these variations by simply expressing the respective PDE's in our computer algebra representation and providing this as an input to our solver. This would allow an exhaustive comparison of all these methods and we consider this as future work.

>>>>

In this paper, we introduced the $P_N$-method to the toolbox of deterministic methods for rendering participating media in computer graphics. We derived and presented the real-valued $P_N$-equations, along with a staggered grid solver for solving them. We showed how to use the results in a rendering system and ran our solver for various standard problems for validation.

@@

The literature in other fields often states that the $P_N$ method---when solved in normal form like we do---is able to deal with vacuum regions. We found this misleading. The $P_N$-method in normal form indeed does not break down completely in the presence of vacuum as diffusion based methods do (due to $\sigma_t^{-1}$ in the diffusion coefficient). However, in the presence of vacuum, the condition number of the system matrix becomes infinite and does not converge either. Therefore $P_N$ based methods also require minimum thresholding of the extinction coefficient and offer no benefit for vacuum regions.

@@

Unique to our system is that it uses a computer algebra representation of the equation to solve as input. Discretization in angular and spatial domain is done using manipulation passes on the representation. The stencil code, which is used to populate the system of linear equations, is generated from the expression tree. This way, we can easily deal with coupled PDEs and avoid the time consuming and error prone process of writing stencil code by hand.

When researching the application of $P_N$ in other fields, we came across a rich variety of variations on the $P_N$-method, such as simplified $P_N$ ($SP_N$)~\cite{Ryan10}, filtered $P_N$ ($FP_N$)~\cite{Radice13}, diffusion-correction $P_N$ ($DP_N$)~\cite{Schaefer11} and least-squares $P_N$ ($LSP_N$)~\cite{Hansen14}. These variations have been introduced to address certain problems of the standard $P_N$-method, such as ringing artifacts, dealing with vacuum regions and general convergence. Our solver can be applied to any (potentially coupled) PDE and therefore can generate stencil code for all these variations by simply expressing the respective PDEs in our computer algebra representation and providing this as an input to our solver. This would allow an exhaustive comparison of all these methods and we consider this as future work.

### 2018-06-25 latex/egsr17-cgf/section_discretized_rte.tex
adding gitignore
applied camera ready fixes

Our method derives from the radiative transfer equation (RTE), which expresses the change of the radiance field $L$, with respect to an infinitesimal change of position into direction $\omega$ at point $\vec{x}$:
\begin{align}
%\label{eq:rte}
\left(\nabla\cdot\omega\right)L\left(\vec{x}, \omega \right)
=&
-\sigma_t\left(\vec{x}\right) L\left(\vec{x}, \omega \right)\nonumber\\
&
+\sigma_s\left(\vec{x}\right) \int_{\Omega}
{
p\left(\omega'\cdot\omega\right)L\left(\vec{x}, \omega \right)\ud\omega'
}\nonumber\\
&
+Q\left(\vec{x}, \omega\right)\nonumber
\quad .
\label{eq:rte}
\end{align}

@@

Deterministic methods are derived by discretizing the angular and spatial domain. This gives rise to a linear system of equations, which can be solved using standard methods. For the $P_N$-method, the angular variable is first discretized, using a truncated spherical harmonics expansion. This results in the $P_N$-equations, a system of coupled PDE's that still depend on a continuous spatial variable.

The number of equations grows with the truncation order $N$. This is why the discretization is hard and error prone if done by hand. We therefore use a computer algebra representation to automate this process. After giving an outline of the general discretization in this section, we will present our computer algebra representation in the next section. The $P_N$-equations that result from our automated discretization are given in Section~\ref{sec:real_valued_pn_eq}.

@@

\begin{figure}[h]
\centering
%\missingfigure{test}
\includegraphics[width=\columnwidth]{figures/fig_matrix_layout.pdf}
\vspace{-0.2in}
\icaption{Structure of coefficient matrix $A$ and solution vector $\vec{u}$ after discretization of the $P_N$-equations on a finite difference grid.}
\label{fig:matrix_layout}
\end{figure}

>>>>

Our method derives from the radiative transfer equation (RTE), which expresses the change of the radiance field $L$, with respect to an infinitesimal change of position in direction $\omega$ at point $\vec{x}$:
\begin{align}
%\label{eq:rte}
\left(\nabla\cdot\omega\right)L\left(\vec{x}, \omega \right)
=&
-\sigma_t\left(\vec{x}\right) L\left(\vec{x}, \omega \right)\nonumber\\
&
+\sigma_s\left(\vec{x}\right) \int_{\Omega}
{
p\left(\omega'\cdot\omega\right)L\left(\vec{x}, \omega \right)\ud\omega'
}\nonumber\\
&
+Q\left(\vec{x}, \omega\right)\nonumber
\quad .
\label{eq:rte}
\end{align}

@@

Deterministic methods are derived by discretizing the angular and spatial domain. This gives rise to a linear system of equations, which can be solved using standard methods. For the $P_N$-method, the angular variable is first discretized, using a truncated spherical harmonics expansion. This results in the $P_N$-equations, a system of coupled PDEs that still depend on a continuous spatial variable.

The number of equations grows with the truncation order $N$. This is why the discretization is laborious and difficult to do without errors if done by hand. We therefore use a computer algebra representation to automate this process. After giving an outline of the general discretization in this section, we will present our computer algebra representation in the next section. The $P_N$-equations that result from our automated discretization are given in Section~\ref{sec:real_valued_pn_eq}.

@@

\begin{figure}[h]
\centering
%\missingfigure{test}
\includegraphics[width=\columnwidth]{figures/fig_matrix_layout.pdf}
%\missingfigure{figures fig matrix layout}
\vspace{-0.2in}
\icaption{Structure of coefficient matrix $A$ and solution vector $\vec{u}$ after discretization of the $P_N$-equations on a finite difference grid.}
\label{fig:matrix_layout}
\end{figure}

### 2018-06-25 latex/egsr17-cgf/section_results.tex
adding gitignore
applied camera ready fixes

\begin{figure}[h]
\centering
\begin{subfigure}{0.49\columnwidth}
%\missingfigure{test}
\includegraphics[width=\columnwidth]{images/checkerboard2d_p1_neumann_staggered_starmap.png}
\end{subfigure}%
\hspace{0.01\columnwidth}
\begin{subfigure}{0.49\columnwidth}
%\missingfigure{test2}
\includegraphics[width=\columnwidth]{images/checkerboard2d_p1_neumann_staggered.png}
\end{subfigure}%
\vspace{-0.2in}
\icaption{Comparison of the result for the checkerboard test using StaRMAP's time-stepping solver~\cite{Seibold14} (left) against our steady-state solver (right) with $N=5$. Our results are in good agreement.}
\label{fig:vs_starmap}
\end{figure}

As can be seen in figure~\ref{fig:vs_starmap}, the results from our solver are in good agreement with the results from Seibold et al.~\cite{Seibold14} and verify the correctness of our implementation. Converging to a residual of $10e^{-10}$ takes $0.27s$ for $P_1$ and $25s$ for $P_5$.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\columnwidth}
\includegraphics[width=\columnwidth]{figures/pointsource_pn.pdf}
%\missingfigure{test}
\caption{$P_N$ vs. ground truth}
\label{fig:pointsource_pn}
\end{subfigure}%
\hspace{0.05\columnwidth}
\begin{subfigure}{0.45\columnwidth}
%\missingfigure{test2}
\includegraphics[width=\columnwidth]{figures/pointsource_p5.pdf}
\caption{$P_5$ vs. CDA and FLD}
\label{fig:pointsource_p5}
\end{subfigure}%
\vspace{-0.1in}
\icaption{Lineplot through the 3D solution of our solver for the point source problem for various order $N$ (left). Solution for $P_5$ compared against classical diffusion, flux-limited diffusion and analytical solution (right).}
\label{fig:pointsource}
\end{figure}

@@

Finally, we run our solver on a procedural cloud dataset to get an idea of its performance in more practical applications. Figure~\ref{fig:teaser}~\subref{fig:nebulae_ours} shows the result of $P_5$ for a procedurally generated heterogeneous cloud with an isotropic phase function. We see that at order $N=5$, our method can capture indirect illumination similarily well as FLD and is significantly better than CDA as expected. The indirectly illuminated region at the bottom appears to be closer to the path-traced result as opposed to the solution from FLD which is very flat in that region. However, in many other areas, $P_5$ seems to still suffer a lot from the problem of energy loss near transitions from dense regions to vacuum. It appears that going higher order a few steps mitigates this only slowly at a high cost of compute and storage.

@@

\begin{figure}[h]
\centering
%\missingfigure{test}
\includegraphics[width=0.9\columnwidth]{figures/fig_nebulae_p1_convergence.pdf}
%\vspace{-0.2in}
\vspace{-0.1in}
\icaption{Convergence behavior of our solver with $N=1$ for the nebulae dataset and for varying minimum thresholds of the extinction coefficient $\sigma_t$. Threshold values and an estimate for the condition number of $A$ (Matlabs condest function) are shown next to the plots. The convergence deteriorates as the threshold decreases. Once it reaches zero, the presence of pure vacuum makes the condition number infinite.}
\label{fig:results_convergence}
\end{figure}

>>>>

\begin{figure}[h]
\centering
\begin{subfigure}{0.49\columnwidth}
\includegraphics[width=\columnwidth]{images/checkerboard2d_p1_neumann_staggered_starmap.png}
\end{subfigure}%
\hspace{0.01\columnwidth}
\begin{subfigure}{0.49\columnwidth}
\includegraphics[width=\columnwidth]{images/checkerboard2d_p1_neumann_staggered.png}
\end{subfigure}%
\vspace{-0.2in}
\icaption{Comparison of the result for the checkerboard test using \textsf{StaRMAP}'s time-stepping solver~\cite{Seibold14} (left) against our steady-state solver (right) with $N=5$. Our results are in good agreement.}
\label{fig:vs_starmap}
\end{figure}

As can be seen in Figure~\ref{fig:vs_starmap}, the results from our solver are in good agreement with the results from Seibold et al.~\cite{Seibold14} and verify the correctness of our implementation. Converging to a residual of $10e^{-10}$ takes $0.27s$ for $P_1$ and $25s$ for $P_5$.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\columnwidth}
\includegraphics[width=\columnwidth]{figures/pointsource_pn.pdf}
%\missingfigure{figures fig pointsource pn}
\caption{$P_N$ vs. ground truth}
\label{fig:pointsource_pn}
\end{subfigure}%
\hspace{0.05\columnwidth}
\begin{subfigure}{0.45\columnwidth}
\includegraphics[width=\columnwidth]{figures/pointsource_p5.pdf}
%\missingfigure{figures fig pointsource p5}
\caption{$P_5$ vs. CDA and FLD}
\label{fig:pointsource_p5}
\end{subfigure}%
\vspace{-0.1in}
\icaption{Lineplot through the 3D solution of our solver for the point source problem for various order $N$ (left). Solution for $P_5$ compared against classical diffusion, flux-limited diffusion and analytical solution (right).}
\label{fig:pointsource}
\end{figure}

@@

Finally, we run our solver on a procedural cloud dataset to get an idea of its performance in more practical applications. Figure~\ref{fig:teaser}\subref{fig:nebulae_ours} shows the result of $P_5$ for a procedurally generated heterogeneous cloud with an isotropic phase function. We see that at order $N=5$, our method can capture indirect illumination similarly well as FLD and is significantly better than CDA as expected. The indirectly illuminated region at the bottom appears to be closer to the path-traced result as opposed to the solution from FLD which is very flat in that region. However, in many other areas, $P_5$ seems to still suffer a lot from the problem of energy loss near transitions from dense regions to vacuum. It appears that going higher order a few steps mitigates this only slowly at a high cost of compute and storage.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.9\columnwidth]{figures/fig_nebulae_p1_convergence.pdf}
%\missingfigure{figures fig nebulae p1 convergence}
\vspace{-0.1in}
\icaption{Convergence behavior of our solver with $N=1$ for the nebulae dataset and for varying minimum thresholds of the extinction coefficient $\sigma_t$. Threshold values and an estimate for the condition number of $A$ (Matlabs condest function) are shown next to the plots. The convergence deteriorates as the threshold decreases. Once it reaches zero, the presence of pure vacuum makes the condition number infinite.}
\label{fig:results_convergence}
\end{figure}
